Sparse Function-space Representation of Neural Networks

### 2023-09-05 workshops/icml-duality/main.tex
updated text and fits on 4 arxiv pages

\begin{abstract}
Deep neural networks (NNs) are known to lack uncertainty estimates, struggle to incorporate new data, and suffer from catastrophic forgetting. We present a method that mitigates these issues by converting NNs from weight-space to a low-rank (sparse) function-space representation, via a dual parameterization. Importantly, our sparse representation captures the joint distribution over the entire data set. This offers a compact and principled way of capturing uncertainty and enables us to incorporate new data without retraining whilst retaining predictive performance. We provide proof-of-concept demonstrations with the proposed approach for quantifying uncertainty in supervised learning on UCI benchmark tasks.
\end{abstract}

\section{Introduction}
%
Deep learning \citep{goodfellow2016deep} has become the cornerstone of contemporary artificial intelligence, proving remarkably effective in tackling supervised and unsupervised learning tasks in the {\em large data}, {\em offline}, and {\em gradient-based training} regime. Despite its success, gradient-based learning techniques exhibit limitations. Firstly, how can we efficiently quantify uncertainty without resorting to expensive and hard-to-interpret sampling in the model's weight-space? Secondly, how to update the weights of an already trained model with new batches of data without compromising the performance on past data?
These questions become central when applied to sequential learning paradigms, such as continual learning \citep[CL,][]{parisi2019continual, de2021continual}, Bayesian optimization \citep[BO,][]{garnett_bayesoptbook_2022} and reinforcement learning  \citep[RL,][]{sutton2018reinforcement}.
% In CL, access to the previous data is lost, and then the challenge is retaining a compact representation of the problem to alleviate forgetting over the life-long learning horizon~\citep{mccloskey1989catastrophic}.
% Similarly, in RL, the model must adapt to environmental observations and can leverage predictive uncertainties for exploration.\looseness-1

@@

%Current state of affairs
Recent techniques  \citep[\eg,][]{ritter2018kfac,khan2019approximate,daxberger2021laplace,fortuin2021bayesian,immer2021scalable} apply a Laplace-GGN approximation to convert trained NNs into Bayesian neural networks (BNNs), that can provide uncertainty without sacrificing additional resources to training \citep{foong2019between}. Furthermore, the resultant weight-space posterior can be converted to the function-space as shown in \citet{khan2019approximate, immer2021improving}. The function-space representation allows for a principled mathematical approach for analyzing its behaviour \citep{cho2009kernel,meronen2020stationary}, performing probabilistic inference \citep{khan2019approximate}, and quantifying uncertainty in NNs \citep{foong2019between}. These methods rely on the linearization of the NN and the resultant neural tangent kernel \citep[NTK,][]{jacot2018neural}.
The NN is characterized in function-space by its first two moment functions, a mean function and covariance function (or kernel)---defining a Gaussian process \citep[GP,][]{rasmussen2006gaussian}. GPs provide a widely used probabilistic toolkit with principled uncertainty estimates.
% They serve as a standard surrogate model for Bayesian optimization \citep{garnett_bayesoptbook_2022} and are effective in model-based reinforcement learning \citep{deisenroth2011pilco} with theoretical guarantees on regret bounds \citep{srinivas2009gaussian}.
%Yet many problems lie in high dimensional input space; for example, images are where GPs cannot learn representations. In such scenarios such as in many reinforcement learning environments neural networks are used as the surrogate model. However, uncertainty is still essential to ensure effective exploration for sequential algorithms. Successful approaches have attempted to blend neural networks with uncertainty estimates around predictions, allowing for sophisticated exploration strategies. However, there has been limited use of hybrid models that possess the feature representation ability of neural networks but also attractive the properties of GPs, such a hybrid method we propose in this paper.

%Need for adaptive learning methods + failures with current methods
Given an approximate inference technique, we demonstrate that NN's emit `dual' parameters which are the building blocks of a GP~\citep{csato2002sparse, adam2021dual, chang2023memory}.
In contrast to previous work that utilizes subsets of training data \citep{immer2021scalable}, our parameterization captures
information from {\em all} data points in a sparse representation, essential for scaling to deep learning data sets.
% predictive uncertainty.
Importantly, the dual parameterization can be used to 1) sparsify the GP without requiring further optimization
(e.g. variational inference), and 2) can incorporate new data without retraining (in a computationally efficient manner
by conditioning on new data).
% In contrast to previous work that utilizes subsets of training data \citep{immer2021scalable}, our parameterization captures information from {\em all} data points in a sparse representation, essential for predictive uncertainty.
% Importantly, our method predicts in the same space as the original trained NN, with the benefit of avoiding the
% complexity introduced by working in weight-space and the notorious cubic complexity of vanilla GPs.
%\todo{Could make it clearer that the GP predicts in the output space while avoiding the NN parameter space}
%, a feature not present in previous approaches, while avoiding the notorious cubic complexity of vanilla GPs.
Through the dual parameterization, we establish a connection between the NN, GPs, and a sparse GP approximations similar to sparse variational GPs~\citep{titsias2009variational}. We refer to our method as Sparse Function-space Representation (\our)---a sparse GP derived from a trained NN; see \cref{fig:teaser} for an example. %Moreover, this dual parameterization can be exploited to perform dual conditioning \citep{chang2022fantasizing}, \ie, an effective approach for conditioning on new data without needing to retrain the model (see \cref{fig:teaser}).
%As \our is in the dual parameters, we can perform dual conditioning recently shown effective in \cite{chang2022fantasizing}, that is, avoid retraining and condition new data into our model (see \cref{fig:teaser}) \todo{Effective compared to what?}.
\looseness-2

@@

\textbf{Bayesian NNs}
In Bayesian deep learning, we place a prior over the weights $p(\vw)$ and aim to calculate the posterior over the weights given the data $p(\vw \mid \mathcal{D})$.
Given the posterior over the weights $p(\vw \mid \mathcal{D})$, we can then make probabilistic predictions with,
\begin{align} \label{eq-predictions}
  p_{\text{\sc BNN}}(y \mid \vx, \mathcal{D}) = \E_{p(\vw \mid \mathcal{D})} \left[ p(y \mid f_{\vw}(\vx)) \right]
\end{align}
The posterior ${p(\vw \mid \dataset) \propto p(\vy \mid f_{\weights}(\vx)) \, p(\weights)}$ is generally not available in closed form
so we resort to approximations.
% Sampling methods that characterize the posterior with a finite set of samples in weight-space are general-purpose, but computationally heavy and impractical for downstream applications.
% In this paper, we build on the Laplace approximation

@@

% \subsection{Laplace approximation}
\textbf{MAP}
It is common to train the NN's weights $\vw$ to minimize the (regularized) empirical risk,
%
\begin{align} \label{eq-empirical-risk}
  \weights^{*} &=
                 \arg \min_{\weights} \mathcal{L}(\dataset,\weights) \nonumber \\
               &=
  \arg \min_{\weights} \textstyle\sum_{i=1}^{N} \underbrace{\ell(f_\weights(\mathbf{x}_{i}), y_i)}_{-\log p(y_{i} \mid f_{\vw}(\vx_{i}))} + \underbrace{\delta \mathcal{R}(\weights)}_{-\log p(\vw)}.
\end{align}
%
This objective corresponds to the log joint distribution $\mathcal{L}(\mathcal{D},\vw)=p(\mathcal{D}, \vw)$
as the loss can be interpreted as a negative log likelihood $\ell(f_\weights(\vx_{i}), y_i) = -\log(p(y_{i} \mid f_\weights(\vx_{i}))$
and the regularizer corresponds to a log prior over the weights $\delta\mathcal{R}(\weights) = \log p(\vw)$.
For example, a weight decay regularizer $\mathcal{R}(\vw) = \frac{1}{2}\|\weights\|^{2}_2$ corresponds to a Gaussinan prior over the weights $p(\vw) = \Norm(\vzeros, \delta^{-1} \MI)$ with
prior precision $\delta$.
As such, we can view \cref{eq-empirical-risk} as the maximum {\it a~posteriori} (MAP) solution.

@@

\textbf{Laplace-GGN}
The Laplace approximation \citet{aa} builds upon this and approximates the weight posterior $p(\vw \mid \mathcal{D}) \approx q(\vw)= \mathcal{N}(\vw^{*} , \bm\Sigma)$
around the MAP weights ($\vw^{*}$) by setting the covariance to the Hessian of the posterior,
\begin{align} \label{eq:laplace-cov}
 \bm\Sigma = - \left[\nabla^{2}_{\vw\vw} \log p(\vw \mid \mathcal{D})  |_{\vw=\vw^{*}} \right]^{-1}.
\end{align}
Computing \cref{eq:laplace-cov} requires calculating the Hessian of the log-likelihood $\nabla^{2}_{\vw\vw} \log p(\vy \mid f_{\vw}(\vx))$ from \cref{eq-empirical-risk}.
As highlighted in \citet{immer2021improving}, computing this Hessian is often infeasible and in practice it is common to adopt the
generalised Gauss-Newton (GGN) approximation,
\begin{align}
 \nabla^{2}_{\vw\vw} \log p(\vy \mid f_{\vw}(\vx)) \approx \Jac{\vw}{\vx}^{\top} \nabla_{\vf\vf}^{2}\log(\vy\mid\vf) \Jac{\vw}{\vx}.
  \nonumber
\end{align}
% \textbf{Laplace-GGN converts BNN into GLM}
\citet{immer2021improving} highlighted that the GGN approximation corresponds to a local linearization of the NN,
% Recent work \citep{khan2019approximate,maddox2021fast} has converted NNs to GPs by linearizing around the NN's MAP weights (obtained from \cref{eq-empirical-risk}).
% Recent work \citep{khan2019approximate,maddox2021fast} has shown that NN (in weight space) leads to a GP, i.e. a function-space approximation.
$f_{\weights^{*}}^{\text{lin}}(\vx) = f_{\weights}(\vx) + \Jac{\weights^{*}}{\vx}(\weights-\weights^{*})$.
This suggests that predictions should be made with a generalised linear model (GLM), given by,
\begin{align} \label{eq-glm}
  p_{\text{\sc GLM}}(\vy \mid \vx, \mathcal{D}) &= \E_{q(\vw)} \left[ p(\vy \mid f_{\vw^{*}}^{\text{lin}}(\vx))\right]. %\nonumber \\
  % &\approx \frac{1}{S} \sum_{s} p(\vy \mid f_{\vw_{s}}^{\text{lin}}(\vx)), \quad \vw_{s} \sim q(\vw)
\end{align}

\textbf{Weight space GLM is equivalent to a GP}
As Gaussian distributions remain tractable under linear transformations, we can covert our model from
weight space to function space \citep[see Ch.~2.1 in ][]{rasmussen2006gaussian}.
 % highlighted that the GGN approximation corresponds to a local linearization of the NN,
As such (and shown in \citet{immer2021improving}), the Bayesian GLM in \cref{eq-glm} has an equivalent GP formulation,
% \begin{align}
%   p_{\text{\sc GP}}(\vy \mid \vx, \mathcal{D}) &= \E_{q(\vf)} \left[ p(\vy \mid \vf)\right] %\nonumber \\
%   % &\approx \frac{1}{S} \sum_{s} p(\vy \mid f_{\vw_{s}}^{\text{lin}}(\vx)), \quad \vw_{s} \sim q(\vw)
% \end{align}
% where the GP posterior is given by,
\begin{subequations}  \label{eq-gp-pred-immer}
\begin{align}
  p_{\text{\sc GP}}(\vy \mid \vx, \mathcal{D}) &= \E_{q(\vf)} \left[ p(\vy \mid \vf)\right] \\ %\nonumber \\
  \myexpect_{p(f_i \mid\vy)}[f_i] &= f_{\vw^{*}}(\vx_{i}), \\
  \mathrm{Var}_{p(f_i \mid \vy)}[f_i] &= k_{ii} - \vk_{\vx i}^\top ( \MKxx + \bm\Lambda^{-1})^{-1} \vk_{\vx i},
\end{align}
\end{subequations}
where the kernel $\kappa(\vx,\vx')=\frac{1}{\delta} \Jac{\weights^*}{\vx} \, \JacT{\weights^*}{\vx'}$ is the Neural Tangent Kernel \citep[NTK,][]{jacot2018neural},
based on the Jacobian at the MAP $\Jac{\weights}{\vx} \coloneqq \left[ \nabla_\weights f_\weights(\vx)\right]^\top \in \R^{1 \times P}$
and $f_i = f_\vw(\vx_i)$ denotes the function output at input $\vx_i$.
The $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{N \times N}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where
each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, $k_{ii} = \kappa(\vx_i, \vx_i)$ and $\bm\Lambda = - \nabla^2_{\vf \vf}\log p(\vy \mid \vf)$ can be interpreted as per-input noise.
% For notational conciseness, we % \cref{eq-laplace-approx-function-space},
% restrict our notatin to a single function output. The extension to multiple outputs is straightforward.

@@

\textbf{Sparse Gaussian processes}
The GP formulation from \citet{immer2021improving} (shown in \cref{eq-gp-pred-immer})
requires inverting an $N\times N$ matrix which has complexity $O(N^3)$.
This limits its applicability to large data sets, which are common in deep learning.
Sparse GPs reduce the computational complexity by representing the GP as a low-rank approximation at a set of inducing
inputs $\vz \in \R^{D}$ with corresponding inducing variables $\vu = f(\vz)$ \citep[see][for an early overview]{quinonero2005unifying}.
The approach by \citet{titsias2009variational} \citep[also used in the DTC approximation, see][]{quinonero2005unifying},
defines the marginal predictive distribution as $q_{\vu}(f_i)  = \int p(f_i  \mid \vu) \, q(\vu) \, \mathrm{d}\vu$ where
the variational distribution is parameterized as $q(\vu) = \mathcal{N}\left(\vu \mid \vm, \MS \right)$ (as in \citet{titsias2009variational,hensman2013gaussian}).
\todo{is this what titsias used or just hensman?}
This gives the sparse GP predictive posterior,
\begin{subequations}  \label{eq-svgp-pred}
\begin{align}
  \myexpect_{q_{\vu}(f_i)}[f_i] &= f_{\vw^{*}}(\vx_{i}) + \vk_{\vz i}^{\top}\MKzz^{-1}(\vm - f_{\vw^{*}}(\vz)), \\
  \mathrm{Var}_{q_{\vu}(f_i)}[f_i] &= k_{ii} - \vk_{\vx i}^\top \MKzz^{-1}( \MKzz - \MS )^{-1} \MKzz^{-1} \vk_{\vx i},  \nonumber
\end{align}
\end{subequations}
where $\MKzz$ and $\vkzs$ are defined similarly to $\MKxx$ and $\vk_{\vx i}$ but over the inducing points $\{\vz_j\}_{i=1}^M=\vz \in \R^{M \times D}$.%, instead of the full data set $\dataset$.
% To the best of our knowledge, there are no methods that convert NNs to sparse GPs.
It is worth noting that the parameters of the variational posterior ($\vm$ and $\MS$) are usually obtained via variational
inference, which requires further optimization.% and is non trivial in the NN setting.

Not only does the GP formulation from \citet{immer2021improving} (shown in \cref{eq-gp-pred-immer}) struggle to scale to large data sets but
it also cannot be incorporate new data by conditioning on it.
This is because it's posterior mean is the NN  $f_{\vw^*}(\vx)$.
Our method overcomes both of these limitations via a dual parameterization.
Importantly, our dual parameterization enables us to 1) sparsify the GP withput resorting to further optimisation, and 2)
incorporate new data without retraining (in a computationally efficient manner).

@@

\paragraph{Dual parameters from NN}
Given that we use a Laplace approximation of the NN, we remove the expectation over the posterior \citep[see Ch.3.4.1 in][for derivation]{rasmussen2006gaussian} and our dual parameters are given by,
%
\begin{subequations}
\label{eq:dual_param_laplace}
\begin{align}
  \hat{\alpha}_i &\coloneqq \nabla_{f}\log p(y_i \mid f) |_{f=f_i} , \\
  %\quad \text{and} \quad
  \hat{\beta}_i &\coloneqq - \nabla^2_{ff}\log p(y_i \mid f) |_{f=f_i}.
  \end{align}
\end{subequations}
%
Substituting \cref{eq:dual_param_laplace} into \cref{eq:gp_pred}, we obtain our GP based on the converged NN. The problem with forming predictions using \cref{eq:gp_pred} is that they cost $O(N^3)$, which limits its applicability to large data sets.

@@

%
We evaluate \our on eight UCI~\citep{UCI} binary and multi-class classification tasks.
Following \citet{immer2021improving}, we used a two-layer MLP with width 50 and tanh activation functions and we used half of the test data set as a validation data set.
We trained the NN using Adam \cite{adam} with a learning rate of $10^{-4}$ and a batch size of $128$.
Training was stopped when the validation NLPD stopped decreasing after 1000 steps.
The checkpoint with lowest NLPD was the NN MAP.
Each experiment was run for $5$ seeds.\looseness-1

>>>>

\begin{abstract}
Deep neural networks (NNs) are known to lack uncertainty estimates and struggle to incorporate new data. We present a method that mitigates these issues by converting NNs from weight space to function space, via a dual parameterization. Importantly, the dual parameterization enables us to formulate a sparse representation that captures information from the entire data set. This offers a compact and principled way of capturing uncertainty and enables us to incorporate new data without retraining whilst retaining predictive performance. We provide proof-of-concept demonstrations with the proposed approach for quantifying uncertainty in supervised learning on UCI benchmark tasks.
\end{abstract}

\section{Introduction}
%
Deep learning \citep{goodfellow2016deep} has become the cornerstone of contemporary artificial intelligence, proving remarkably effective in tackling supervised and unsupervised learning tasks in the {\em large data}, {\em offline}, and {\em gradient-based training} regime. Despite its success, gradient-based learning techniques exhibit limitations. Firstly, how can we efficiently quantify uncertainty without resorting to expensive and hard-to-interpret sampling in the model's weight space? Secondly, how to update the weights of an already trained model with new batches of data without compromising the performance on past data?
These questions become central when used for sequential learning,
such as continual learning \citep[CL,][]{parisi2019continual, de2021continual}, Bayesian optimization \citep[BO,][]{garnett_bayesoptbook_2022} and reinforcement learning  \citep[RL,][]{sutton2018reinforcement}.
% In CL, access to the previous data is lost, and then the challenge is retaining a compact representation of the problem to alleviate forgetting over the life-long learning horizon~\citep{mccloskey1989catastrophic}.
% Similarly, in RL, the model must adapt to environmental observations and can leverage predictive uncertainties for exploration.\looseness-1

@@

%Current state of affairs
Recent techniques  \citep[\eg,][]{ritter2018kfac,khan2019approximate,daxberger2021laplace,fortuin2021bayesian,immer2021scalable} apply a Laplace
generalized Gauss-Newton (GGN) approximation to convert trained NNs into Bayesian neural networks (BNNs), that can provide uncertainty without sacrificing additional resources to training \citep{foong2019between}. Furthermore, the resultant weight-space posterior can be converted to function space as shown in \citet{khan2019approximate, immer2021improving}. The function-space representation allows for a principled mathematical approach for analyzing its behaviour \citep{cho2009kernel,meronen2020stationary}, performing probabilistic inference \citep{khan2019approximate}, and quantifying uncertainty \citep{foong2019between}. These methods rely on linearizing the NN and the resultant neural tangent kernel \citep[NTK,][]{jacot2018neural}.
The NN is characterized in function space by its first two moment functions, a mean function and covariance function (or kernel)---defining a Gaussian process \citep[GP,][]{rasmussen2006gaussian}. GPs provide a widely used probabilistic toolkit with principled uncertainty estimates.
% They serve as a standard surrogate model for Bayesian optimization \citep{garnett_bayesoptbook_2022} and are effective in model-based reinforcement learning \citep{deisenroth2011pilco} with theoretical guarantees on regret bounds \citep{srinivas2009gaussian}.
%Yet many problems lie in high dimensional input space; for example, images are where GPs cannot learn representations. In such scenarios such as in many reinforcement learning environments neural networks are used as the surrogate model. However, uncertainty is still essential to ensure effective exploration for sequential algorithms. Successful approaches have attempted to blend neural networks with uncertainty estimates around predictions, allowing for sophisticated exploration strategies. However, there has been limited use of hybrid models that possess the feature representation ability of neural networks but also attractive the properties of GPs, such a hybrid method we propose in this paper.

%Need for adaptive learning methods + failures with current methods
Given an approximate inference technique, we demonstrate that NNs emit `dual' parameters which are the building blocks of a GP~\citep{csato2002sparse, adam2021dual, chang2023memory}.
In contrast to previous work that utilizes subsets of training data \citep{immer2021scalable}, our parameterization captures
information from {\em all} data points in a sparse representation, essential for scaling to deep learning data sets.
% predictive uncertainty.
Importantly, the dual parameterization can be used to 1) sparsify the GP without requiring further optimization
(e.g. variational inference) whilst capturing information from all data points, and 2) can incorporate new data without retraining (in a computationally efficient manner
by conditioning on new data).
% In contrast to previous work that utilizes subsets of training data \citep{immer2021scalable}, our parameterization captures information from {\em all} data points in a sparse representation, essential for predictive uncertainty.
% Importantly, our method predicts in the same space as the original trained NN, with the benefit of avoiding the
% complexity introduced by working in weight-space and the notorious cubic complexity of vanilla GPs.
%\todo{Could make it clearer that the GP predicts in the output space while avoiding the NN parameter space}
%, a feature not present in previous approaches, while avoiding the notorious cubic complexity of vanilla GPs.
Through the dual parameterization, we establish a connection between the NN, GPs, and sparse approximations similar to sparse variational GPs~\citep{titsias2009variational}. We refer to our method as Sparse Function-space Representation (\our)---a sparse GP derived from a trained NN; see \cref{fig:teaser} for an example. %Moreover, this dual parameterization can be exploited to perform dual conditioning \citep{chang2022fantasizing}, \ie, an effective approach for conditioning on new data without needing to retrain the model (see \cref{fig:teaser}).
%As \our is in the dual parameters, we can perform dual conditioning recently shown effective in \cite{chang2022fantasizing}, that is, avoid retraining and condition new data into our model (see \cref{fig:teaser}) \todo{Effective compared to what?}.
\looseness-2

@@

% \textbf{Bayesian NNs}
\textbf{BNNs}
In Bayesian deep learning, we place a prior over the weights $p(\vw)$ and aim to calculate the posterior over the weights given the data $p(\vw \mid \mathcal{D})$.
Given the weight posterior $p(\vw \mid \mathcal{D})$, we can make probabilistic predictions with,
\begin{align} \label{eq-predictions}
  p_{\text{\sc BNN}}(y \mid \vx, \mathcal{D}) = \E_{p(\vw \mid \mathcal{D})} \left[ p(y \mid f_{\vw}(\vx)) \right]
\end{align}
The posterior ${p(\vw \mid \dataset) \propto p(\vy \mid f_{\weights}(\vx)) \, p(\weights)}$ is generally not available in closed form
so we resort to approximations.
% Sampling methods that characterize the posterior with a finite set of samples in weight-space are general-purpose, but computationally heavy and impractical for downstream applications.
% In this paper, we build on the Laplace approximation

@@

% \subsection{Laplace approximation}
\textbf{MAP}
It is common to train NN weights $\vw$ to minimize the (regularized) empirical risk,
%
\begin{align} \label{eq-empirical-risk}
  \weights^{*} &=
                 \arg \min_{\weights} \mathcal{L}(\dataset,\weights) \nonumber \\
               &=
  \arg \min_{\weights} \textstyle\sum_{i=1}^{N} \underbrace{\ell(f_\weights(\mathbf{x}_{i}), y_i)}_{-\log p(y_{i} \mid f_{\vw}(\vx_{i}))} + \underbrace{\delta \mathcal{R}(\weights)}_{-\log p(\vw)}.
\end{align}
%
This objective corresponds to the log joint distribution $\mathcal{L}(\mathcal{D},\vw)=p(\mathcal{D}, \vw)$
as the loss can be interpreted as a negative log likelihood $\ell(f_\weights(\vx_{i}), y_i) = -\log(p(y_{i} \mid f_\weights(\vx_{i}))$
and the regularizer corresponds to a log prior over the weights $\delta\mathcal{R}(\weights) = -\log p(\vw)$.
For example, a weight decay regularizer $\mathcal{R}(\vw) = \frac{1}{2}\|\weights\|^{2}_2$ corresponds to a Gaussian prior over the weights $p(\vw) = \Norm(\vzeros, \delta^{-1} \MI)$, with
prior precision $\delta$.
As such, we can view \cref{eq-empirical-risk} as the maximum {\it a~posteriori} (MAP) solution.

@@

\textbf{Laplace approximation}
The Laplace approximation \citet{mackayBayesian1992,daxberger2021laplace} builds upon this and approximates the weight posterior $p(\vw \mid \mathcal{D}) \approx q(\vw)= \mathcal{N}(\vw^{*} , \bm\Sigma)$
around the MAP weights ($\vw^{*}$) by setting the covariance to the Hessian of the posterior,
\begin{align} \label{eq:laplace-cov}
 \bm\Sigma = - \left[\nabla^{2}_{\vw\vw} \log p(\vw \mid \mathcal{D})  |_{\vw=\vw^{*}} \right]^{-1}.
\end{align}
Computing \cref{eq:laplace-cov} requires calculating the Hessian of the log-likelihood $\nabla^{2}_{\vw\vw} \log p(\vy \mid f_{\vw}(\vx))$ from \cref{eq-empirical-risk}.
As highlighted in \citet{immer2021improving}, computing this Hessian is often infeasible and in practice it is common to adopt the
generalized Gauss-Newton (GGN) approximation,
\begin{align}
 \nabla^{2}_{\vw\vw} \log p(\vy \mid f_{\vw}(\vx)) \approx \Jac{\vw}{\vx}^{\top} \nabla_{\vf\vf}^{2}\log(\vy\mid\vf) \Jac{\vw}{\vx}.
  \nonumber
\end{align}
where the Jacobian is given by $\Jac{\weights}{\vx} \coloneqq \left[ \nabla_\weights f_\weights(\vx)\right]^\top \in \R^{1 \times P}$.
% \textbf{Laplace-GGN converts BNN into GLM}
\citet{immer2021improving} highlighted that the GGN approximation corresponds to a local linearization of the NN,
% Recent work \citep{khan2019approximate,maddox2021fast} has converted NNs to GPs by linearizing around the NN's MAP weights (obtained from \cref{eq-empirical-risk}).
% Recent work \citep{khan2019approximate,maddox2021fast} has shown that NN (in weight space) leads to a GP, i.e. a function-space approximation.
$f_{\weights^{*}}^{\text{lin}}(\vx) = f_{\weights}(\vx) + \Jac{\weights^{*}}{\vx}(\weights-\weights^{*})$.
This suggests that predictions should be made with a generalized linear model (GLM), given by,
\begin{align} \label{eq-glm}
  p_{\text{\sc GLM}}(y_{i} \mid \vx, \mathcal{D}) &= \E_{q(\vw)} \left[ p(y_{i} \mid f_{\vw^{*}}^{\text{lin}}(\vx))\right]. %\nonumber \\
  % &\approx \frac{1}{S} \sum_{s} p(\vy \mid f_{\vw_{s}}^{\text{lin}}(\vx)), \quad \vw_{s} \sim q(\vw)
\end{align}

% \textbf{Weight space GLM is equivalent to a GP}
\textbf{Gaussian processes}
As Gaussian distributions remain tractable under linear transformations, we can covert the linear model from
weight space to function space \citep[see Ch.~2.1 in ][]{rasmussen2006gaussian}.
 % highlighted that the GGN approximation corresponds to a local linearization of the NN,
As such (and shown in \citet{immer2021improving}), the Bayesian GLM in \cref{eq-glm} has an equivalent GP formulation,
% \begin{align}
%   p_{\text{\sc GP}}(\vy \mid \vx, \mathcal{D}) &= \E_{q(\vf)} \left[ p(\vy \mid \vf)\right] %\nonumber \\
%   % &\approx \frac{1}{S} \sum_{s} p(\vy \mid f_{\vw_{s}}^{\text{lin}}(\vx)), \quad \vw_{s} \sim q(\vw)
% \end{align}
% where the GP posterior is given by,
% \begin{subequations}  \label{eq-gp-pred-immer}
% \begin{align}
%   p_{\text{\sc GP}}(\vy \mid \vx, \mathcal{D}) &= \E_{q(\vf)} \left[ p(\vy \mid \vf)\right] \\ %\nonumber \\
%   \myexpect_{p(f_i \mid\vy)}[f_i] &= f_{\vw^{*}}(\vx_{i}), \\
%   \mathrm{Var}_{p(f_i \mid \vy)}[f_i] &= k_{ii} - \vk_{\vx i}^\top ( \MKxx + \bm\Lambda^{-1})^{-1} \vk_{\vx i},
% \end{align}
% \end{subequations}
\begin{align} \label{eq-gp-pred-immer}
  p_{\text{\sc GP}}(y_{i} \mid \vx, \mathcal{D}) &= \E_{q(f_{i})} \left[ p(y_{i} \mid f_{i} )\right]
\end{align}
where $q(f_{i}) = \mathcal{N} \left( f_{\vw^{*}}(\vx_{i}), \sigma^{2}_{i} \right)$ is the GP's predictive posterior with variance,
% where $q(\vf) = \mathcal{N} \left( \myexpect_{p(f_i \mid\vy)}[f_i], \mathrm{Var}_{p(f_i \mid \vy)}[f_i] \right)$ is the predictive GP posterior given by,
\begin{align} \label{eq-gp-pred-immer-gp}
  % \myexpect_{p(f_i \mid\vy)}[f_i] &= f_{\vw^{*}}(\vx_{i}), \\
  \sigma^{2}_{i} &= k_{ii} - \vk_{\vx i}^\top ( \MKxx + \bm\Lambda^{-1})^{-1} \vk_{\vx i},
\end{align}
where the kernel $\kappa(\vx,\vx')=\frac{1}{\delta} \Jac{\weights^*}{\vx} \, \JacT{\weights^*}{\vx'}$ is the Neural Tangent Kernel \citep[NTK,][]{jacot2018neural}
and $f_i = f_\vw(\vx_i)$ is the function output at $\vx_i$.
The $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{N \times N}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where
each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, $k_{ii} = \kappa(\vx_i, \vx_i)$ and $\bm\Lambda = - \nabla^2_{\vf \vf}\log p(\vy \mid \vf)$ can be interpreted as per-input noise.
% For notational conciseness, we % \cref{eq-laplace-approx-function-space},
% restrict our notatin to a single function output. The extension to multiple outputs is straightforward.

@@

\textbf{Sparse Gaussian processes}
The GP formulation from \citet{immer2021improving} (shown in \cref{eq-gp-pred-immer})
requires inverting an $N\times N$ matrix which has complexity $O(N^3)$.
This limits its applicability to large data sets, which are common in deep learning.
Sparse GPs reduce the computational complexity by representing the GP as a low-rank approximation at a set of inducing
inputs $\vz \in \R^{D}$ with corresponding inducing variables $\vu = f(\vz)$ \citep[see][for an early overview]{quinonero2005unifying}.
The approach by \citet{titsias2009variational} \citep[also used in the DTC approximation, see][]{quinonero2005unifying},
defines the marginal predictive distribution as $q_{\vu}(f_i)  = \int p(f_i  \mid \vu) \, q(\vu) \, \mathrm{d}\vu$ where
the variational distribution is parameterized as $q(\vu) = \mathcal{N}\left(\vu \mid \vm, \MS \right)$ (as in \citet{titsias2009variational,hensman2013gaussian}).
This gives the sparse GP predictive posterior,
\begin{subequations}  \label{eq-svgp-pred}
\begin{align}
  \myexpect_{q_{\vu}(f_i)}[f_i] &= f_{\vw^{*}}(\vx_{i}) + \vk_{\vz i}^{\top}\MKzz^{-1}(\vm - f_{\vw^{*}}(\vz)), \\
  \mathrm{Var}_{q_{\vu}(f_i)}[f_i] &= k_{ii} - \vk_{\vx i}^\top ( \MKzz -  \MKzz^{-1}\MS  \MKzz^{-1}))^{-1} \vk_{\vx i},  \nonumber
\end{align}
\end{subequations}
where $\MKzz$ and $\vkzs$ are defined similarly to $\MKxx$ and $\vk_{\vx i}$ but over the inducing points $\{\vz_j\}_{i=1}^M=\vz \in \R^{M \times D}$.%, instead of the full data set $\dataset$.
% To the best of our knowledge, there are no methods that convert NNs to sparse GPs.
It is worth noting that the parameters of the variational posterior ($\vm$ and $\MS$) are usually obtained via variational
inference, which requires further optimization.% and is non trivial in the NN setting.

Not only does the GP formulation from \citet{immer2021improving} (shown in \cref{eq-gp-pred-immer}) struggle to scale to large
data sets but it also cannot incorporate new data by conditioning on it.
This is because it's posterior mean is the NN  $f_{\vw^*}(\vx)$.
Our method overcomes both of these limitations via a dual parameterization.
Importantly, our dual parameterization enables us to 1) sparsify the GP without further optimization, and 2)
incorporate new data without retraining.\looseness-1 % (in a computationally efficient manner).

@@

\paragraph{Dual parameters from NN}
Given that we use a Laplace approximation of the NN, we remove the expectation over the posterior \citep[see Ch.3.4.1 in][for derivation]{rasmussen2006gaussian} and our dual parameters are given by,
%
\begin{subequations}
\label{eq:dual_param_laplace}
\begin{align}
  \hat{\alpha}_i &\coloneqq \nabla_{f}\log p(y_i \mid f) |_{f=f_i} , \\
  %\quad \text{and} \quad
  \hat{\beta}_i &\coloneqq - \nabla^2_{ff}\log p(y_i \mid f) |_{f=f_i}.
  \end{align}
\end{subequations}
%
Substituting \cref{eq:dual_param_laplace} into \cref{eq:gp_pred}, we obtain our GP based on the trained NN.
Making predictions with \cref{eq:gp_pred} costs $O(N^3)$, which limits its applicability on large data sets.

@@

%
We evaluate \our on eight UCI~\citep{UCI} binary and multi-class classification tasks.
% Following \citet{immer2021improving}, we used a two-layer MLP with width 50 and tanh activation functions and we used half of the test data set as a validation data set.
Following \citet{immer2021improving}, we used a two-layer MLP with width 50, tanh activation functions and a
$70\%\ (\text{train}):15\%\ (\text{validation}) :15\%\ (\text{test})$ data split.
We trained the NN using Adam \cite{adam} with a learning rate of $10^{-4}$ and a batch size of $128$.
Training was stopped when the validation NLPD stopped decreasing after 1000 steps.
The checkpoint with lowest NLPD was the NN MAP.
Each experiment was run for $5$ seeds.\looseness-1

% We first compare \our to other posthoc Bayesian deep learning approaches.
% In particular, we compare to the Laplace approximation when making predictions with 1) the nonlinear NN (BNN) and 2) the generalised linear model (GLM).
We first compare \our to the Laplace approximation when making predictions with 1) the nonlinear NN (BNN) and 2) the generalised linear model (GLM) in \cref{eq-glm}.
We use the Laplace PyTorch library \citep{daxberger2021laplace} with the full Hessian.%, i.e. we do not resort to the diagonal or KFAC approximations.
It is common practice to tune the prior precision $\delta$ after training the NN.
That is, find a prior precision $\delta_{\text{tune}}$ which has a better NLPD on the validation set than the $\delta$ used to train the NN.
Note that this leads to different prior precisions being used for training and inference.
We highlighted that technically this invalidates the Laplace/SFR methods as the NN weights are no longer the MAP weights.
Nevertheless, we report results when the prior precision $\delta$ is tuned (\cref{tbl:uci} right) and is not tuned (\cref{tbl:uci} left).
When tuning the prior precision $\delta$, \our matches the performance of the Laplace methods (BNN/GLM).
Interestingly, \our also  performs well without tuning the prior precision $\delta$, unlike the Laplace methods (BNN/GLM).
